Comparing to lognormal, Poisson, exponential

### 2021-03-09 analysis/rev_prototyping/ct_mcdb.Rmd
use median

---
title: "Shannon diversity"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(grid)

@@

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20)

```
```{r}

ggplot(all_di, aes(nsingletons_percentile)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat, singletons), scales = "free_y")

ggplot(all_di, aes(nsingletons_percentile_excl)) +
  geom_histogram(bins = 100) +
  facet_wrap(vars(dat, singletons), scales = "free_y")

all_di <- all_di %>%
  mutate(prop_singles_actual = nsingletons / s0,
         sim_prop_singles_95 = nsingletons_95 / s0,
         sim_prop_singles_mean = nsingletons_mean / s0) %>%
  mutate(actual_minus_95_prop = prop_singles_actual - sim_prop_singles_95) %>%
  mutate(high_singles=nsingletons_percentile > 95,
         high_singles_ex = nsingletons_percentile_excl > 95,
         low_singles = nsingletons_percentile < 5,
         low_singles_ex = nsingletons_percentile_excl < 5) %>%
  mutate(singles_desc = ifelse(high_singles_ex, "high", ifelse(low_singles, "low", "int")))

@@

```{r, fig.dim = c(8,8)}

ggplot(all_di, aes(s0, n0, color = singles_desc)) +
  geom_point(alpha = .3) +
  facet_wrap(vars(dat), scales = "free") +
  scale_x_log10()+ 
  scale_y_log10()

all_di %>%
  group_by(dat, singletons) %>%
  summarize(prop_high_singles = mean(high_singles),
            prop_high_singles_ex = mean(high_singles_ex),
            prop_low_singles = mean(low_singles),
            prop_low_singles_ex = mean(low_singles_ex),
            nsites= dplyr::n()) %>%
  ungroup()

```{r, echo = T}

ggplot(all_di, aes(n0/s0, nsingletons_percentile_excl, color = singles_desc)) +
  geom_point() +
  facet_wrap(vars(dat), scales = "free") +
  scale_x_log10() +
  geom_vline(xintercept = c(3, 5, 10))

ggplot(all_di, aes(n0/s0, nsingletons_percentile_excl, color = singles_desc)) +
  geom_point(data = filter(all_di, singles_desc != "int")) +
  geom_point(alpha = .1, color = "grey") +
  facet_wrap(vars(dat), scales = "free") +
  scale_x_log10() +
  geom_vline(xintercept = c(3, 5, 10))

all_di %>%
  filter(n0/s0 < 3) %>%
  group_by(dat, singletons) %>%
  summarize(prop_high_singles = mean(high_singles),
            prop_high_singles_ex = mean(high_singles_ex),
            prop_low_singles = mean(low_singles),
            prop_low_singles_ex = mean(low_singles_ex),
            nsites= dplyr::n()) %>%
  ungroup()

all_di %>%
  filter(n0/s0 > 10) %>%
  group_by(dat, singletons) %>%
  summarize(prop_high_singles = mean(high_singles),
            prop_high_singles_ex = mean(high_singles_ex),
            prop_low_singles = mean(low_singles),
            prop_low_singles_ex = mean(low_singles_ex),
            nsites= dplyr::n()) %>%
  ungroup()

```
```{r}

all_di <- all_di %>%
  mutate(singletons_change = (nsingletons - nsingletons_95) / s0)

ggplot(filter(all_di, high_singles_ex), aes(singletons_change)) +
  geom_histogram() +
  facet_wrap(vars(dat), scales = "free")

all_di %>%
  filter(high_singles_ex) %>%
  group_by(dat) %>%
  summarize(mean_singles_change = mean(singletons_change),
            n = dplyr::n())

>>>>

---
title: "Comparing to lognormal, Poisson, exponential"
author: Renata Diaz
date: "`r Sys.Date()`"
output: 
    github_document:
       df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsanalysis)

knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, message = FALSE)

@@

mcdb_ct <- read.csv(here::here("mcdb_cts.csv"))

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log_s0 = log(s0),
         log_n0 = log(n0)) %>%
  filter(n0 != s0,
         s0 != 1,
         !singletons,
         n0 != (s0 + 1)) %>%
  mutate(dat = ifelse(grepl(dat, pattern = "fia"), "fia", dat),
         dat = ifelse(dat == "misc_abund_short", "misc_abund", dat)) %>%
  mutate(Dataset = dat,
         Dataset = ifelse(Dataset == "fia", "FIA", Dataset),
         Dataset = ifelse(Dataset == "bbs", "Breeding Bird Survey", Dataset),
         Dataset = ifelse(Dataset == "mcdb", "Mammal Communities", Dataset),
         Dataset = ifelse(Dataset == "gentry", "Gentry", Dataset),
         Dataset = ifelse(Dataset == "misc_abund", "Misc. Abundance", Dataset)) %>%
  filter(nparts > 20) %>%
  filter(dat == "mcdb") %>%
  left_join(mcdb_ct)
```

```{r}

ggplot(all_di, aes(sim_pos_from_best)) +
  geom_histogram()

ggplot(all_di, aes(sim_pos_from_best, real_po)) +
  geom_point() +
  geom_line(aes(sim_pos_from_best, sim_pos_from_best))

@@
